---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# The Johnson Quantile-Parameterised Distribution

<!-- badges: start -->
<!-- badges: end -->

The goal of `rjqpd` is to implement the Johnson Quantile-Parameterised 
distribution system in R. It provides the density function, cumulative 
distribution function, quantile function and random sampling function in their 
usual form:

* `djqpd` - calculates the density function (PDF)
* `pjqpd` - calculates the cumulative distribution function (CDF)
* `qjqpd` - calculates the quantile function (QF)
* `rjqpd` - generates random samples

## Background

A common problem in data analysis is encountering continuous uncertainties that 
are not easily characterised by well known parametric, continuous probability distributions. Often the underlying process that generated the data is unknown
or is inconvenient to express in closed form. In these cases one typically has
to resort to curve fitting a distribution to a set of probability-quantile pairs
without the guarantee that the best-fit distribution will pass through the
assessed points exactly.

The Johnson Quantile-Parameterised Distribution (J-QPD) system is a novel 
approach to this problem first developed in a [paper](https://pubsonline.informs.org/doi/10.1287/deca.2016.0343) by Hadlock 
and Bickel (2017) in Decision Analysis Vol. 14, No. 1.

Their approach is to extend the [Johnson Distribution System](https://en.wikipedia.org/wiki/Johnson's_SU-distribution) in a way such
that the distributions constructed honour exactly a symmetric percentile triplet of
quantile assessments (e.g. 15th, 50th, 85th) on a bounded or semi-bounded domain.
By parameterising the distribution in this way the need for curve fitting can be
eliminated.

## Installation

The `rjqpd` package is not yet available on CRAN.

You can install the latest development version of `rjqpd` using:

``` r
devtools::install_github("bobbyingram/rjqpd")
```

## Basic Usage

The base function of the package is the `jqpd` class constructor:

```{r, eval = FALSE}
jqpd()
```

This function takes the convenient symmetric percentile triplet quantile parameterisation and computes a set of constants required to calculate the 
distribution functions.

## Basic Example

This is a basic example to show how to calculate the parameters of a bounded
Johnson Quantile-Parameterised distribution from the quantiles of a symmetric 
percentile triplet (SPT).

```{r basic example}
library(rjqpd)

# alpha = 0.1 indicates a symmetric percentile triplet (SPT) of (0.1, 0.5, 0.9)
alpha <- 0.1
# we parameterise using some quantile values corresponding to the SPT above.
quantiles <- c(0.32, 0.40, 0.60)
params <- jqpd(quantiles, lower = 0, upper = 1, alpha = alpha)

# The CDF should return the (SPT) at the quantile values:
pjqpd(quantiles, params)
```

### Summary Plot

The `plot_jqpd` function produces a simple summary plot showing the density,
cumulative distribution function, quantile function and 1000 random samples:

```{r summary_plot}
plot_jqpd(params)
```

### Summary Statistics

We also provide some basic summary statistics functions:

* `jqpd_mean` - calculates the mean
* `jqpd_var` - calculates the variance
* `jqpd_sd` - calculates the standard deviation
* `jqpd_skewness` - calculates the skewness
* `jqpd_kurtosis` - calculates the kurtosis

```{r}
jqpd_mean(params)
```

```{r}
jqpd_var(params)
```

```{r}
jqpd_sd(params)
```

```{r}
jqpd_skewness(params)
```

```{r}
jqpd_kurtosis(params)
```
